\ttitle

### 2013-05-31 Chapters/Chapter2.tex
Ch2. Sec. CUSUM
Work on Ch2

%----------------------------------------------------------------------------------------
\section{Outline}
\begin{itemize}
  \item Literature review about Temporal Segmentation (previous draft was more about classification)
  \item Consider methods for the context of filter-methods for classification
  \item Take a loot at 3-4 different kind of methods for change detection:
    \begin{itemize}
      \item Dimensionality reduction
      \item Density-ratio estimation
      \item Support Vector Machines (?) - if there are more sources about this
      \item CUSUM - or other more traditional methods
    \end{itemize}
  \item With each method, shortly look at characteristics, strengths and weaknesses and consider applicability to accelerometer sensor data
\end{itemize}

@@

The widely used CUSUM (cumulative sum) method by Basseville \etal \cite{basseville1993detection}, and the GLR (Generalized Likelihood Ratio) by Gustafsson \cite{gustafsson1996marginalized,gustafsson2000adaptive} take this approach.
The former originates from control methods for detection from bench marks.
The latter compares the logarithm of the likelihood ratio over two consecutive intervals.
These two methods is discussed and analyzed in section \ref{cusum-glr}.

@@

The final method under consideration, change detection via (intrinsic) dimensionality reduction, takes a different point of view.
Opposed to the other discussed methods, dimensionality reduction is framed in the MDL (Minimum Description Length) framework.
It uses the estimated underlying number of parameters of the time series as a model for change detection.
Section \ref{dim-reduction} discuses this method.

%----------------------------------------------------------------------------------------
% CUSUM and GLR
\input{./Chapters/Chapter2/cusum_glr}

@@

%----------------------------------------------------------------------------------------
% Dimensionality reduciton
\input{./Chapters/Chapter2/dim_reduction}

>>>>

%----------------------------------------------------------------------------------------
\section{Outline}
\begin{itemize}
  \item Literature review about Temporal Segmentation (previous draft was more about classification)
  \item Consider methods for the context of filter-methods for classification
  \item Take a loot at 3-4 different kind of methods for change detection:
    \begin{itemize}
      \item Introduction with a lot of techniques
      \item Explain why look at a few
      \item CUSUM - or other more traditional methods
      \item Density-ratio estimation
      \item Support Vector Machines (?) - if there are more sources about this
      \item (Dimensionality reduction)
    \end{itemize}
  \item With each method, shortly look at characteristics, strengths and weaknesses and consider applicability to accelerometer sensor data
\end{itemize}

@@

*** Remove notion of GLR ***

The widely used CUSUM (cumulative sum) method by Basseville \etal \cite{basseville1993detection}, and the GLR (Generalized Likelihood Ratio) by Gustafsson \cite{gustafsson1996marginalized,gustafsson2000adaptive} take this approach.
The former originates from control methods for detection from bench marks.
The latter compares the logarithm of the likelihood ratio over two consecutive intervals.
These two methods is discussed and analyzed in section \ref{cusum}.

@@

% The final method under consideration, change detection via (intrinsic) dimensionality reduction, takes a different point of view.
% Opposed to the other discussed methods, dimensionality reduction is framed in the MDL (Minimum Description Length) framework.
% It uses the estimated underlying number of parameters of the time series as a model for change detection.
% Section \ref{dim-reduction} discuses this method.

%----------------------------------------------------------------------------------------
% CUSUM and GLR
\input{./Chapters/Chapter2/cusum}

@@

%----------------------------------------------------------------------------------------
% Dimensionality reduciton
% \input{./Chapters/Chapter2/dim_reduction}

### 2013-05-31 Chapters/Chapter2/cusum.tex
Work on Ch2

An other used approach in the statistical framework of change detection is the Cumulative Sum of Squares (CUSUM) as introduced by Page \cite{page1954continuous}.
Originally used for quality control in production environments, its main function is to detect change in the mean of measurements.
Many extensions to this method have been proposed.
Some focus on the change in mean, such the method of Alippi and Roveri \cite{alippi2006adaptive}.
Others apply the method the problems in which the change of variance is under consideration.
Among others are there the centered version of the cumulative sums, introduced by Brown, Durbin and Evans \cite{brown1975techniques} and the Moving Sum (MOSUM) of squares by \cite{hsu2007mosum}.

>>>>

An other used approach in the statistical framework of change detection is the Cumulative Sum of Squares (CUSUM) as introduced by Page \cite{page1954continuous}.
Originally used for quality control in production environments, its main function is to detect change in the mean of measurements and has been applied to this problem \cite{basseville1993detection}.
Many extensions to this method have been proposed.
Some focus on the change in mean, such the method of Alippi and Roveri \cite{alippi2006adaptive}.
Others apply the method the problems in which the change of variance is under consideration.
Among others are there the centered version of the cumulative sums, introduced by Brown, Durbin and Evans \cite{brown1975techniques} and the Moving Sum (MOSUM) of squares by \cite{hsu2007mosum}.

### 2013-05-31 Chapters/Chapter2/density_ratio.tex
Work on Ch2

Papers:
\begin{itemize}
  \item Change-Point Detection in Time-Series Data by Direct Density-Ratio Estimation, 2009, 45 refs \cite{kawahara2009change}
  \item Change-Point Detection in Time-Series Data by Relative Density-Ratio Estimation, 2013, 3 refs \cite{liu2013change}
  \item Density ratio estimation in machine learning (Book), 2009, 24 refs \cite{sugiyama2012density}
\end{itemize}

Formulate the problem of detecting change in the statistical framework.
Consider the probability distributions from which two consecutive segments of time series around a target time point are generated.
When the disitrubtions differ significantly the target time point is regarded as a change point.

CUSUM (cumulative sum) \cite{basseville1993detection} and GLR (generalized likelihood ratio)

The distribution over the values of time series data can be represented with a probability density function (pdf).
Two sections of a time series data can be generated with the same underlying pdf or each with a different.

>>>>

Papers:
\begin{itemize}
  \item Change-Point Detection in Time-Series Data by Direct Density-Ratio Estimation, 2009, 45 refs \cite{kawahara2009change}
  \item Change-Point Detection in Time-Series Data by Relative Density-Ratio Estimation, 2013, 3 refs \cite{liu2013change}
  \item Density ratio estimation in machine learning (Book), 2009, 24 refs \cite{sugiyama2012density}
  \item Direct importance estimation for covariate shift adaptation, 2008, 84 refs \cite{sugiyama2008direct}
\end{itemize}

% Formulate the problem of detecting change in the statistical framework.
% Consider the probability distributions from which two consecutive segments of time series around a target time point are generated.
% When the disitrubtions differ significantly the target time point is regarded as a change point.

% CUSUM (cumulative sum) \cite{basseville1993detection} and GLR (generalized likelihood ratio)

% The distribution over the values of time series data can be represented with a probability density function (pdf).
% Two sections of a time series data can be generated with the same underlying pdf or each with a different.

Many approaches to detect change points, regarded as a change in the underlying probabilistic generation, monitor the logarithm of the likelihood ratio between two consecutive intervals.
Some methods which rely on this are novelty detection, maximum-likelihood estimation and online learning of autoregressive models \cite{kawahara2009change}.
A limitation of these methods is that they rely on pre-specified parametric models.
Non-parametric models for density estimation have been proposed, but it is said to be a hard problem.
A solutions to this is to estimate the \emph{ratio} of probabilities instead of the probabilities themselves.
One of the recent methods to achieve this is the Kullback-Leibler Importance Estimation Procedure (KLIEP) by Sugiyama \etal \cite{sugiyama2008direct}.

The method proposed by Kawahara and Sugiyama \cite{kawahara2009change} is composed of an online version of the KLIEP algorithm.
The method also considers \emph{sequences} of samples (rather than samples directly) because the time series samples are generally not independent over time.
An advantage over other non-parametric approaches, such as sequential one-class support vector machines is that the model has an natural cross-validation procedure.
This makes that the value of tuning parameters, such as the kernel bandwidth, can be objectively obtained.

In their formulation change is detected by monitoring the logarithm of the likelihood ratio between the reference (past) and test (current) time intervals
\begin{equation}
  S = \sum_{i=1}^{n_\text{te}} \text{ln} \frac{p_{\text{te}}(\matrixsym{Y}_\text{te}(i))}{p_\text{rf}(\matrixsym{Y}_\text{te}(i))}
\end{equation}
Where $\matrixsym{Y}_\text{te}(i)$ is a sequence of samples from the test interval.
A change is detected when $S > \mu$, for some predetermined threshold $\mu$.
The question is then to how calculate the density ratio
\begin{equation}
  w(\matrixsym{Y}) := \frac{p_\text{te}(\matrixsym{Y)})}{p_\text{rf}(\matrixsym{Y})}
\end{equation}
because this ratio is unknown and should be estimated.
The naive approach is to estimate the ratio by taking the ratio of the estimated densities.
Since this is known to be a hard problem and sensitive for errors, the solution would be to estimate the ratio directly.

The procedure of the method proposed by Kawahara and Sugiyama \cite{kawahara2009change} is to first apply the batch KLIEP algorithm with model selection for initial parameter $\vectorsym{\alpha}$ and kernel width calculation.
Then for every new sample the reference and test intervals are shifted and the calculated parameters $\vectorsym{\alpha}$ are updated.
Finally the logarithm of the likelihood ratio is evaluated.
If it is beyond the threshold $\mu$ the current time is reported as a change point.

Improvements in this line of research, by Liu \etal \cite{liu2013change} has lead the application of improved density-ratio estimation methods to the problem of change detection.
Such an improvement is the Unconstrained Least-Squares Importance Fitting (uLSIF) method \cite{kanamori2009least} and an extension which possess a superior non-parametric convergence property: Relative uLSIF (RuLSIF) \cite{yamada2013relative}.

### 2013-05-31 Chapters/Chapter2/svm.tex
Work on Ch2

Introduced by Vapnik \cite{vapnik1998statistical, vapnik1999nature}, Support Vector Machines offer a way to segment, and classify, linear separable data.
When combined with a mapping function, which maps the data from the input space $I$ to a higher dimension feature space $F$, the input data can be non-linear separable.
The linear hyperplane, which segments the data in the feature space $F$, yields to a non-linear segmentation in the lower-dimensional input space $I$.
Instead of explicitly mapping the input data to the higher dimensional space, a kernel function can be used.
This kernel function can calculate values of the feature space directly, without the need to first map the input values to this space.
This process is referred to as the kernel trick.

*** Let sigma be a mapping from I to F such that the dot product in F can be computed using some simple kernel ***

\subsection{One-class Support Vector Machine}

@@

Using this one-class SVM formulation, differences between two (consecutive) windows of data points with size $w$ can be obtained.
The first window is used as the input set, $h_1$ and the second as the test set $h_t$.
For the first window a one-class SVM is constructed, yielding in a representation by $c_1$ and $r_1$.
When the data points of the second window belong to the same class, the representation of that one-class SVM would equal the first:
\begin{equation}
  c_1 = c_2, r_1 = r_2
\end{equation}

>>>>

Introduced by Vapnik \cite{vapnik1998statistical, vapnik1999nature}, Support Vector Machines offer a way to segment, and classify, linear separable data.
This technique can also be applied to estimate density functions of given time series \cite{weston1999support}.
When combined with a mapping function, which maps the data from the input space $I$ to a higher dimension feature space $F$, the input data can be non-linear separable.
The linear hyperplane, which segments the data in the feature space $F$, yields to a non-linear segmentation in the lower-dimensional input space $I$.
Instead of explicitly mapping the input data to the higher dimensional space, a kernel function can be used.
This kernel function can calculate values of the feature space directly, without the need to first map the input values to this space.
This process is referred to as the kernel trick.

The method of Camci \cite{camci2010change} uses one-class support vector machines to segment time series data.
For the current interval a hyper-sphere is constructed, using the kernel trick, which circumscribes all the data.
To cope with possible errors or outliers a soft-margin is applied.
The data can thus be represented using a center $c_1$ and radius $r_1$ of the hyper-sphere.
New data points are consecutively mapped from the input space to the feature space.
The retrieved (high dimension) data point can be in- or outside the earlier constructed hyper-sphere, thereby giving information about a possible change point.

\subsection{One-class Support Vector Machine}
*** This section is too detailed for the Literature Review, but keep it for now ***

@@

Using this one-class SVM formulation, differences between two (consecutive) windows of data points with size $w$ can be obtained.
The first window is used as the input set, $h_1$ and the second as the test set $h_t$.
For the first window a one-class SVM is constructed, yielding in a representation by $c_1$ and $r_1$.
When the data points of the second window belong to the same class, the representation of that one-class SVM would equal the first:
\begin{equation}
\begin{aligned}
  c_1 = c_2, & &  r_1 = r_2
\end{aligned}
\end{equation}
